---
title: "Time Series Project 2"
author: "Ziping_Wei"
date: "12/20/2019"
output:
  word_document: default
  pdf_document: default
---

```{r}
library(astsa) 
library(tseries) 
library(TSA) 
library(forecast) 
library(dplyr) 
library(lubridate) 
library(ggplot2)
```

1.load the data
source: https://fred.stlouisfed.org/series/MRTSSM4481USN
```{r}
df_orginal = read.csv('cloth09.csv')
#change the column names
names(df_orginal) = c('Date','Sales')
#get the value of sales
df = ts(df_orginal$Sales,frequency = 12, start = c(2009,1))
# train test split
train_ratio = 0.8
splitpoint = floor(length(df)*train_ratio)
train = ts(df[1:splitpoint],frequency = 12, start = c(2009,1)) #2009.01.01-2017.08.01
test = ts(df[(splitpoint+1):length(df)],frequency = 12, start = c(2017,9)) #2017.08.01-2009.10.01

cat(' Total observations:',length(df),
    '\n','Train Observations:',length(train),
    '\n','Test Observations:',length(test),'\n')
```

2.Exploratory Data Analysis
```{r}
#plot the orginal sales
plot.ts(train, xlab = "Year", ylab = "Sales (Millions of Dollars)", main = "U.S. Retail Sales: Clothing Stores") #non-stationary
acf(train, lag.max = 48,main="ACF (orginal data)", xlab = "Lag (1 means 12 months)")
pacf(train, lag.max = 48,main = "PACF (orginal data)",xlab = "Lag (1 means 12 months)")

```
From the plot, we could see that the data is not stationary with an obvious positive linear trend. In 2008, the retail sales shows a slight decline which may beacused of the recession. The ACF and PACF plot futher prove the nonstationarity of the data, with ACF decays very slowly. The ACF also indicates a seasonal pattern of the data, as the ACF spike at lage 12,24,36. To remove notstationarity, I first take one lag differencing.

```{r}
# ndiffs(train)
# take difference: lag=1
train_l1 = diff(train)
plot.ts(train_l1, xlab = "Year", ylab = "U.S. Retail Sales: Clothing Stores (Millions of Dollars)", main = "Time Series Ordinary difference") 
acf(train_l1, lag.max = 48,main="ACF Ordinary difference", xlab = "Lag")
pacf(train_l1, lag.max = 48,main = "PACF Ordinary difference",xlab = "Lag")

```
After taking one lag difference, the time series plot has a constant mean around 0 and a more stable variance comparied to the orginal one. The ACF still fails to tail off at multiples of seasonal period 12, indicating a searsonal differencing of lag 12 needed to achieve stationarity.
# ARIMA(0,1,1)，ARIMA(0,1,2)
# ARIMA(4,1,1)，ARIMA(6,1,2)

```{r}
# take difference of train_l1: lag = 12
train_l1.12 = diff(diff(train_l1, lag = 12))
plot.ts(train_l1.12, xlab = "Year", ylab = "U.S. Retail Sales: Clothing Stores (Millions of Dollars)", main = "Differenced Time Series Ordinary and Seasonal Difference") 
acf(train_l1.12, lag.max = 48,main="ACF Ordinary and Seasonal Difference", xlab = "Lag")
pacf(train_l1.12, lag.max = 48,main = "PACF Ordinary and Seasonal Difference",xlab = "Lag")

```

Candidate models:

# ARIMA(0,1,1)，ARIMA(0,1,2)
# ARIMA(4,1,1)，ARIMA(6,1,2)

ARIMA(0,1,1)x(0,1,1)[12]
ARIMA(0,1,1)x(0,1,4)[12]
ARIMA(0,1,2)x(0,1,1)[12]
ARIMA(0,1,2)x(0,1,4)[12]

Fit the model and compare the AIC and BIC scores
```{r}

SARIMA_models = function(ts,ps,ds,qs,PS,DS,QS,period){
  for (p in ps){
    for (q in qs){
      for (Q in QS){
        order = c(p,ds,q)
        seasonal_order = c(PS,DS,Q)
        model = arima(ts,order = order, 
                      seasonal = list(order=seasonal_order,period = period)
                      # method="CSS-ML",
                      # optim.control = list(maxit = 500)
                      )
        print('====================================')
        print(sprintf("SARIMA(%d,%d,%d)x(%d,%d,%d)[%d]",p,ds,q,PS,DS,Q,period))
        print('====================================')
        print(sprintf('AIC:%f',AIC(model)))
        print(sprintf('BIC:%f',BIC(model)))    
      }
    }
  }
}
```


```{r}
ps = c(0,4,6)
qs = c(1,2)
QS = c(1,4)
SARIMA_models(train,ps,1,qs,0,1,QS,12)

```

Dignostic checking 
AIC: SARIMA(4,1,1)x(0,1,1)[12]

```{r}

sarima(train,4,1,1,0,1,1,12)
```
BIC: SARIMA(0,1,1)x(0,1,1)[12]
```{r}
sarima(train,0,1,1,0,1,1,12)
```

```{r}
arima.4.1.1.0.1.1 = arima(train, order=c(4,1,1),
                          seasonal = list(order = c(0,1,1), period = 12))
acf(arima.4.1.1.0.1.1$residuals, lag.max = 48, main = 'ACF of Residuals')
pacf(arima.4.1.1.0.1.1$residuals, lag.max = 48, main = 'PACF of Residuals')

```

```{r}
#011
arima.0.1.1.0.1.1 = arima(train, order=c(0,1,1),
                          seasonal = list(order = c(0,1,1), period = 12))
acf(arima.0.1.1.0.1.1$residuals, lag.max = 48, main = 'ACF of Residuals')
pacf(arima.0.1.1.0.1.1$residuals, lag.max = 48, main = 'PACF of Residuals')
```

```{r}
yhat_test411 = predict(arima.4.1.1.0.1.1,n.ahead=length(test))
yhat_test411_ts = ts(yhat_test411$pred,frequency = 12, start = c(2017,9)) 
autoplot(df,series='Actual Sales') +
  autolayer(yhat_test411_ts,series='Predicted Sales') +
  labs(y='Sales (Millions of Dollars)', 
       x='Year',
       title = 'U.S. 2009-2019 Retail Sales: Clothing Stores',
       subtitle = 'SARIMA(4,1,1)x(0,1,1)[12]'
       )

```

```{r}
yhat_test011 = predict(arima.0.1.1.0.1.1,n.ahead=length(test))
yhat_test011_ts = ts(yhat_test011$pred,frequency = 12, start = c(2017,9)) 
autoplot(df,series='Actual Sales') +
  autolayer(yhat_test011_ts,series='Predicted Sales') +
  labs(y='Sales (Millions of Dollars)', 
       x='Year',
       title = 'U.S. 2009-2019 Retail Sales: Clothing Stores',
       subtitle = 'SARIMA(0,1,1)x(0,1,1)[12]'
       )

```
```{r}
accuracy(yhat_test411_ts,test)
accuracy(yhat_test011_ts,test)
```


```{r}
####### FINAL ONE#####
# SARIMA(0,1,1)x(0,1,1)[12] trained on the whole dataset
sarima_011 = arima(df, order=c(0,1,1),
                          seasonal = list(order = c(0,1,1), period = 12))
#predict following 1yrs sales
pred_future011 = predict(sarima_011,n.ahead=1*12)
pred_future011_ts = ts(pred_future011$pred,frequency = 12,start = c(2019,11))

autoplot(df,series='Actual Sales') +
  autolayer(pred_future011_ts,series='predicted sales') +
  labs(y='Millions of dollars', 
       x='Year',
       title = 'U.S. 2020 Retail Sales Predication: Clothing Stores',
       subtitle = 'SARIMA(0,1,1)x(0,1,1)[12]')

```


```{r}
# SARIMA(4,1,1)x(0,1,1)[12] trained on the whole dataset
sarima_411 = arima(train, order=c(4,1,1),
                          seasonal = list(order = c(0,1,1), period = 12))
#predict following 1yrs sales - 12months
pred_future411 = predict(sarima_411,n.ahead=1*12)
pred_future411_ts = ts(pred_future411$pred,frequency = 12,start = c(2019,10))

autoplot(df,series='Actual Sales') +
  autolayer(pred_future411_ts,series='predicted sales') +
  labs(y='Millions of dollars', 
       x='Year',
       title = 'U.S. 2020 Retail Sales Predication: Clothing Stores',
       subtitle = 'SARIMA(4,1,1)x(0,1,1)[12]')
```

